---
title: "Plotting PMC Trend"
output: html_notebook
---


```{r, include=FALSE}
library(ggplot2)
#setwd('C:\\Users\\kebisu\\Documents\\Research\\PMCoarse_TLBWCA\\Data')
#setwd('C:\\Users\\ebike\\OneDrive\\Documents\\Research\\PMC_LBW_CA\\Data')
setwd('F:\\Research\\PMC_LBW_CA\\PMC_LBW_CA\\Data')
load('PMC_DF_20170602.RData')#PMC_DF
#setwd('C:\\Users\\kebisu\\Documents\\Research\\PMCoarse_TLBWCA\\Data')
load('PMC_DF_20170602.RData')#PMC_DF
```
For now, ignore extreme values.
Make Plot for Year-Month Trend by using Whole California Data
```{r, echo=FALSE}
DF=aggregate(PMC_Value~Date,mean,data=PMC_DF)
DF$MonthDay_num=as.numeric(format(DF$Date,'%j'))
DF$MonthName=factor(months(DF$Date,abbreviate=TRUE))
DF$Period=cut(as.numeric(substr(DF$Date,1,4)),c(2001,2005,2009,2014),
              labels=c('2001-2004','2005-2008','2009-2013'),right=FALSE,include.highest=TRUE)
```
#ggPlot
```{r}
ggplot(DF,aes(Date,PMC_Value)) +
     geom_point(na.rm=TRUE)+
     theme_bw() #This option is my preference; I like black/white plot
```


Let's put some trend line
```{r}
ggplot(DF,aes(Date,PMC_Value)) +
     geom_point(na.rm=TRUE)+
     theme_bw()+
     geom_smooth(span=0.05,method='loess',size=1.5)+
	coord_cartesian(ylim=c(0,50))+
     theme_bw()
```
* Span number is degree of somoothness. The larger, the smoother. 
* There is some seasonal trend, but still difficult to see how much PMC improves (or getting worse?) over decade.
     + Overall, it is improving.  

Now Let's make a more fancy plot

#ggplot with smooth line
```{r}
head(DF) #Check 
Plot=ggplot(data=DF,aes_string(x="MonthDay_num",y="PMC_Value",col="Period"))+
	geom_smooth(span=0.4,method='loess',size=1.5,se=FALSE)+
     theme_bw()
Plot
```

#Some cosmetic modification
```{r}
title=substitute(paste(PM[10-2.5],' Trend from 1990 to 2015 in US'))
Y_title=substitute(paste(PM[10-2.5],' Value'))
Legend=data.frame(DF[!duplicated(DF$MonthName),c('MonthName','MonthDay_num')])

Plot2=Plot+
     scale_x_continuous('Month',breaks=Legend$MonthDay_num,labels=Legend$MonthName,expand=c(0,0))+
	scale_y_continuous(Y_title)+
	ggtitle(title)+
	coord_cartesian(ylim=c(0,35))+
     theme(plot.title=element_text(hjust = 0.5))
Plot2
```

* Adding points (not recommended)
It is quite difficult to see this pattern from scatter plot
```{r}
Plot2+geom_point()
```

# Issue regarding character/number/date
```{r, include=FALSE}
test=PMC_DF[substr(PMC_DF$FIPS_C,1,5)=='06037',]
test$YM=paste(substr(test$Date,1,4),substr(test$Date,6,7),sep='')
test2=aggregate(PMC_Value~YM+FIPS_C,mean,data=test)
#In this output, YM is defined as character
str(test2$YM)
#Let's plot as it is
ggplot(test2,aes(YM,PMC_Value)) +
     geom_point(na.rm=TRUE)+
     theme_bw()+
     geom_smooth(span=0.05,method='loess',size=1.5)+
	coord_cartesian(ylim=c(0,50))+
     theme_bw()
```
* Output look OK, but x-axis is treated as categorical; not continuous variable. I think this is why they don't generate smooth line

* Now let's change to numeric
```{r, include=FALSE}
test2$YM=as.numeric(test2$YM)
ggplot(test2,aes(YM,PMC_Value)) +
     geom_point(na.rm=TRUE)+
     theme_bw()+
     geom_smooth(span=0.05,method='loess',size=1.5)+
	coord_cartesian(ylim=c(0,50))+
     theme_bw()
```
* Now smoothline is plotted, but they treated year-month as continuous, and output is not what we want.

# solution
* Put day (15th) to YM value and define it as date 
```{r}
test2$YM2=as.Date(paste(substr(test2$YM,1,4),'-',substr(test2$YM,5,6),'-15',sep=''),format='%Y-%m-%d')
head(test2)

ggplot(test2,aes(YM2,PMC_Value)) +
     geom_point(na.rm=TRUE)+
     theme_bw()+
     geom_smooth(span=0.05,method='loess',size=1.5)+
	coord_cartesian(ylim=c(0,50))+
     theme_bw()

```

# Let's make a area plot
```{r}
library(dplyr)
library(reshape2)
YMData=PMC_DF %>%
          filter(substr(FIPS_C,1,5)=='06075') %>%
          group_by(substr(Date,1,7)) %>%
          summarize(PMC=mean(PMC_Value),PM25=mean(PM25_Value),PM10=mean(PM10_Value))
names(YMData)[1]=c('Date')
YMData=mutate(YMData,YM=as.Date(paste(substr(Date,1,4),'-',substr(Date,6,7),'-15',sep=''),format='%Y-%m-%d')) %>%
     select(-Date)
YMData2=melt(YMData,id.vars='YM') %>%
          filter(variable!='PM10')

ggplot(YMData2,aes(YM,value))+
     geom_area(aes(colour=variable,fill=variable), position = 'stack')+
     theme_bw()+
	ggtitle('Trend')+
	coord_cartesian(ylim=c(0,60))+
     theme(plot.title=element_text(hjust = 0.5))

```

```{r}
library(dplyr)
library(reshape2)
YMData=PMC_DF %>%
          filter(substr(FIPS_C,1,5)=='06029') %>%
          group_by(substr(Date,1,4)) %>%
          summarize(PMC=mean(PMC_Value),PM25=mean(PM25_Value),PM10=mean(PM10_Value))
names(YMData)[1]=c('Date')
YMData=mutate(YMData,YM=as.Date(paste(substr(Date,1,4),'-06-15',sep=''),format='%Y-%m-%d')) %>%
     select(-Date)
YMData2=melt(YMData,id.vars='YM') %>%
          filter(variable!='PM10')

ggplot(YMData2,aes(YM,value))+
     geom_area(aes(colour=variable,fill=variable), position = 'stack')+
     theme_bw()+
	ggtitle('Trend')+
	coord_cartesian(ylim=c(0,40))+
     theme(plot.title=element_text(hjust = 0.5))
```
```{r}
RatioDF=PMC_DF %>%
    filter(PM25_Value>0&PMC_Value>0) %>%
    mutate(PM25_PCT=PM25_Value*100/PM10_Value,PMC_PCT=PMC_Value*100/PM10_Value) %>%
    filter(substr(FIPS_C,1,5)=='06029') %>%
    group_by(substr(Date,1,4)) %>%
    summarize(PMC=mean(PMC_PCT),PM25=mean(PM25_PCT))
names(RatioDF)[1]=c('Date')
RatioDF2=mutate(RatioDF,YM=as.Date(paste(substr(Date,1,4),'-06-15',sep=''),format='%Y-%m-%d')) %>%
    select(-Date) %>%
    melt(id.vars='YM')

ggplot(RatioDF2,aes(YM,value))+
     geom_area(aes(colour=variable,fill=variable), position = 'stack')+
     theme_bw()+
  	ggtitle('Trend')+
  	coord_cartesian(ylim=c(0,100))+
     theme(plot.title=element_text(hjust = 0.5))
```

```{r}
temp=PMC_DF %>% 
  filter(substr(FIPS_C,1,5)=='06111') %>%
  mutate(year=substr(Date,1,4)) %>% 
  group_by(year) %>% 
  summarize(PM10=mean(PM10_Value),PM25=mean(PM25_Value),PMC=mean(PMC_Value))

ref_value=temp[1,2] %>% 
  as.numeric()

temp2=temp %>% 
    mutate(PM25_PCT=PM25*100/ref_value,PMC_PCT=PMC*100/ref_value) %>% 
    select(year,PM25_PCT,PMC_PCT) %>% 
    melt(id.vars='year') %>%
    mutate(YM=as.Date(paste(year,'-06-15',sep=''),format='%Y-%m-%d'))

ggplot(temp2,aes(YM,value))+
     geom_area(aes(colour=variable,fill=variable), position = 'stack')+
     theme_bw()+
  	ggtitle('Trend')+
  	coord_cartesian(ylim=c(0,100))+
     theme(plot.title=element_text(hjust = 0.5))
```